Associahedra for finite type cluster algebras \\ and minimal relations between $\b{g}$-vectors

### 2019-04-10 thirdPaper/draft/typeConeAssociahedra.tex
added mixed volume open question

\smallskip
\item mixed volumes
\end{itemize}

@@

\begin{example}
\label{exm:typeCone}
\enlargethispage{.5cm}
Consider the $2$-dimensional fan~$\Fan$ depicted in \cref{fig:typeCone} (left) (we will encounter this fan again in \cref{exm:typeConeCoarsenings,exm:typeConeMinkowskiSum,exm:typeConeAsso}).
It has five rays labeled~$1, \dots, 5$ and five maximal cones labeled~$a,\dots, e$.
Its type cone~$\ctypeCone(\Fan)$ lies in~$\R^5$, but has a $2$-dimensional linearity space.
The five pairs of adjacent maximal cones of~$\Fan$ give rise to following five defining inequalities for~$\ctypeCone(\Fan)$:
\begin{gather*}
H^>_{ab}: \; \b{h}_1 + \b{h}_3 - \b{h}_2 > 0
\qquad
H^>_{bc}: \; \b{h}_2 + \b{h}_4 - \b{h}_3 > 0
\qquad
H^>_{cd}: \; \b{h}_3 + \b{h}_5 - \b{h}_4 > 0
\\
H^>_{de}: \; \b{h}_1 + \b{h}_4 > 0
\qquad
H^>_{ae}: \; \b{h}_2 + \b{h}_5 > 0,
\end{gather*}
where $H^>_{ij}$ denotes the halfspace defined by the inequality corresponding to the two adjacent maximal cones~$i$ and~$j$.
Note that the inequalities~$H^>_{ae}$ and~$H^>_{de}$ are redundant.
For example, the height vector~$(\nicefrac{1}{2}, \nicefrac{3}{4}, \nicefrac{5}{4}, 1, \nicefrac{5}{4})$ belongs to the type cone~$\ctypeCone(\Fan)$, and the corresponding polytope is represented in \cref{fig:typeCone} (middle).
To represent this type cone, it suffices to represent~$\b{K}\ctypeCone(\Fan)$ which is an essential simplicial cone in~$\R^3$.
One can further reduce the dimension by intersecting with a transversal hyperplane to get a $2$-dimensional representation.
This is the red triangle depicted in \cref{fig:typeCone} (right).

\begin{figure}[htpb]
	\capstart
	\centerline{{\includegraphics[scale=1]{typeConeFan}} \quad {\includegraphics[scale=1]{typeConePolytope}} \quad \raisebox{.5cm}{\includegraphics[scale=.45]{typeCone}}}
	\caption{A $2$-dimensional fan~$\Fan$ with five rays $1, \dots, 5$ and five maximal cones~$a, \dots, e$ (left), its polytopal realization corresponding to the height vector~$(\nicefrac{1}{2}, \nicefrac{3}{4}, 1, 1, \nicefrac{5}{4})$ (middle), and the intersection of $\b{K}\ctypeCone(\Fan)$ with a hyperplane to get a $2$-dimensional slice (right).}
	\label{fig:typeCone}
\end{figure}
\end{example}

@@

In other words, extremal adjacent pairs define the extremal rays of the polar of the closed type cone~$\ctypeCone(\Fan)$.
Understanding the extremal adjacent pairs of~$\Fan$ enables to describe its type cone~$\typeCone(\Fan)$ and thus all its polytopal realizations.
For instance in

>>>>

\smallskip
\item Once we get a representation of a polytope~$P$ as a signed Minkowski sum~$P=\sum_{1\leq i\leq k}\alpha_i Q_i$, we get a formula for the volume of $P$ as a multivariate polynomial in $\alpha_1,\dots,\alpha_k$ whose coefficients are the mixed volumes of $Q_1,\dots, Q_k$~\cite{ArdilaBenedettiDoker,McMullen-Valuations}. These encode rich combinatorial information when representing generalized permuahedra as positive Minkowski sum of coordinate simplices or hypersimplices~\cite{Postnikov}, and in particular for matroid polytopes~\cite{ArdilaBenedettiDoker}. It would be interesting to get such volume formulas for polytopal realizations of $\b{g}$-vector fans of cluster algebras, and combinatorial interpretations for their coefficients.
\end{itemize}

@@

\begin{example}
\label{exm:typeCone}
\enlargethispage{.5cm}
Consider the $2$-dimensional fan~$\Fan$ depicted in \cref{fig:typeCone} (left) (we will encounter this fan again in \cref{exm:typeConeCoarsenings,exm:typeConeMinkowskiSum,exm:typeConeAsso}).
It has five rays labeled~$1, \dots, 5$ and five maximal cones labeled~$a,\dots, e$.
Its type cone~$\typeCone(\Fan)$ lies in~$\R^5$, but has a $2$-dimensional linearity space.
The five pairs of adjacent maximal cones of~$\Fan$ give rise to following five defining inequalities for~$\typeCone(\Fan)$:
\begin{gather*}
H^>_{ab}: \; \b{h}_1 + \b{h}_3 - \b{h}_2 > 0
\qquad
H^>_{bc}: \; \b{h}_2 + \b{h}_4 - \b{h}_3 > 0
\qquad
H^>_{cd}: \; \b{h}_3 + \b{h}_5 - \b{h}_4 > 0
\\
H^>_{de}: \; \b{h}_1 + \b{h}_4 > 0
\qquad
H^>_{ae}: \; \b{h}_2 + \b{h}_5 > 0,
\end{gather*}
where $H^>_{ij}$ denotes the halfspace defined by the inequality corresponding to the two adjacent maximal cones~$i$ and~$j$.
Note that the inequalities~$H^>_{ae}$ and~$H^>_{de}$ are redundant.
For example, the height vector~$(\nicefrac{1}{2}, \nicefrac{3}{4}, \nicefrac{5}{4}, 1, \nicefrac{5}{4})$ belongs to the type cone~$\typeCone(\Fan)$, and the corresponding polytope is represented in \cref{fig:typeCone} (middle).
To represent this type cone, it suffices to represent~$\b{K}\typeCone(\Fan)$ which is an essential simplicial cone in~$\R^3$.
One can further reduce the dimension by intersecting with a transversal hyperplane to get a $2$-dimensional representation.
This is the red triangle depicted in \cref{fig:typeCone} (right).

\begin{figure}[htpb]
	\capstart
	\centerline{{\includegraphics[scale=1]{typeConeFan}} \quad {\includegraphics[scale=1]{typeConePolytope}} \quad \raisebox{.5cm}{\includegraphics[scale=.45]{typeCone}}}
	\caption{A $2$-dimensional fan~$\Fan$ with five rays $1, \dots, 5$ and five maximal cones~$a, \dots, e$ (left), its polytopal realization corresponding to the height vector~$(\nicefrac{1}{2}, \nicefrac{3}{4}, 1, 1, \nicefrac{5}{4})$ (middle), and the intersection of $\b{K}\typeCone(\Fan)$ with a hyperplane to get a $2$-dimensional slice (right).}
	\label{fig:typeCone}
\end{figure}
\end{example}

@@

In other words, extremal adjacent pairs define the extremal rays of the polar of the closed type cone~$\ctypeCone(\Fan)$.
Understanding the extremal adjacent pairs of~$\Fan$ enables to describe its type cone~$\typeCone(\Fan)$ and thus all its polytopal realizations.
For instance in 
\arnau{unfinished sentence}
